---
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r Packages, include=FALSE}
require(ggplot2)
require(tidyr)
require(dplyr)
require(grid)
require(gridExtra)
```

EXECUTIVE SUMMARY
```{r}
#The whole point of an executive summary is to compress a lengthy report in a few sentences so that a person with very limited time to invest (like an executive director) would be able to understand what the whole thing is about. It should include all: the problem, the methods used and the results obtained.
```
This report aims at investigating the following questions on the base of the database mtcars:

1. Is an automatic or manual transmission better for MPG (miles per gallons)     
2. Quantify the MPG difference between automatic and manual transmissions     

The database contains `r length(mtcars$am)` car models of which `r table(mtcars$am)[1]` have automatic transmission and `r table(mtcars$am)[2]` have manual.

```{r Dataset}
data(mtcars)
mtcars$cyl <- factor(mtcars$cyl); mtcars$vs <- factor(mtcars$vs) 
mtcars$am <- factor(mtcars$am); mtcars$gear <- factor(mtcars$gear) 
mtcars$carb <- factor(mtcars$carb)
head(mtcars)
```

```{r Exploratory graph creation, fig.width=9, include=FALSE}
names_var <- names(select(mtcars, -mpg))
out <- NULL
g <- ggplot(mtcars, aes(y = mpg))
for(i in 1:length(names_var)){
    
    g <- g + aes_string(x = names_var[i]) + geom_point(aes(colour = factor(am))) + geom_smooth(method = "lm", aes(group = 1)) + guides(colour = FALSE)
    
    out[[i]] <- g # creates a list with all the plots 
}
grid.arrange(out[[1]], out[[2]], out[[3]], out[[4]], out[[5]], out[[6]], out[[7]], out[[8]], out[[9]], out[[10]], nrow = 2)
```
    
First we check the null hypothesis that manual and automatic transmissions cars are from the same population (the variable "am" is 0 for automatic transmission and 1 for manual.). For the sake of this projects we assume "mpg" to have a normal distribution, although it looks a little skewed. 
```{r T.test, include = FALSE}
t <- t.test(mtcars$mpg ~ mtcars$am)
```
A simple t-test confirm that the two transmissions are not from the same population as the interval does not contain 0 and the p-value `r t$p.value` is sufficiently low.
```{r t.test interval}
t$conf.int
```

We consider a linear model with "am" as a predictor and "mpg" as a response.    
   
```{r Fit model 1, fig.height= 3, fig.width= 4}
out[[8]]
fit1 <- lm(mpg ~ am -1, mtcars)
summary(fit1)
```
The relatevly high F-statistic and the very low p-value confirm that there might be a correlation, but if we look at how well the model fit, R-squared `r round(summary(fit1)$r.squared, 3)` is very low (1 being best fit, 0 being no fit at all). This, together with a Residual Standard Error (RSE) of `r round(sd(summary(fit1)$residuals), 3)`, suggests that the model is not very accurate and that there might be other model to better explain the correlation.
We look therefore to other variables that might be confounder. 

```{r Fit model 2}
fit2 <- lm(mpg ~ am + cyl + disp + hp + drat +wt + qsec + vs + gear + carb -1, mtcars)
summary(fit2)
```

The only variable with a significant impact and a low p-value is "wt", the weight.

```{r}
fit3 <- lm(mpg ~ am + wt-1, mtcars)
plot(mtcars$am, mtcars$mpg)
abline(fit3)
summary(fit3)
summary(lm(mpg ~ am*wt, mtcars))
```






APPENDIX
Exploratory graphs

```{r, fig.width=9, cache=TRUE, results=TRUE, echo=FALSE}
names_var <- names(select(mtcars, -mpg))
out <- NULL
g <- ggplot(mtcars, aes(y = mpg))
for(i in 1:length(names_var)){
    
    g <- g + aes_string(x = names_var[i]) + geom_point(aes(colour = factor(am))) + geom_smooth(method = "lm", aes(group = 1)) + guides(colour = FALSE)
    
    out[[i]] <- g # creates a list with all the plots 
}
grid.arrange(out[[1]], out[[2]], out[[3]], out[[4]], out[[5]], out[[6]], out[[7]], out[[8]], out[[9]], out[[10]], nrow = 2)
```